---
title: "Group 3 and 4 analysis"
author: "Kent Riemondy RBI"
date: "9/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libs, message=FALSE, warning=FALSE, echo=FALSE}
source("../R/utils.R")
```



```{r}
so <- readRDS(file.path("objects", "all_neoplastic.rds"))
```


```{r}
gp_cells <- colnames(so)[!so$subgroup %in% c("SHH", "WNT")]

so <- so[, gp_cells]

so <- FindVariableFeatures(
  so,
  selection.method = "vst",
  nfeatures = 3000,
  verbose = FALSE
)

so <- ScaleData(so, features = VariableFeatures(so), 
                  vars.to.regress = "percent.mt", 
                  verbose = TRUE)

so <- RunPCA(so, 
             features = VariableFeatures(so),
             npcs = 50, 
             verbose = FALSE)

ElbowPlot(so, ndims = 100)

# make graphs and use graphs for UMAP
so <- FindNeighbors(so, 
                      reduction = "pca", 
                      dims = 1:50, 
                      k.param = 50L)


so <- RunUMAP(so, 
              dims = 1:50, 
              n.neighbors = 30L,
              min.dist = 0.4)

so <- FindClusters(so, 
                   resolution = c(0.1, 0.3, 0.5))

so$coarse_clusters_gp34 <- so$RNA_snn_res.0.1

Idents(so) <- "coarse_clusters_gp34"

plot_umap(so, "coarse_clusters_gp34")
plot_umap(so, "UPN")
plot_umap(so, "subtype")
```

## Harmonize

```{r}
library(harmony)
so <- RunHarmony(so,
                 "UPN", 
                 theta = 1.5,
                 plot_convergence = TRUE)

so <- RunUMAP(so, 
              reduction = "harmony", 
              dims = 1:40, 
              min.dist = 0.2,
              n.neighbors = 30L,
              reduction.name = "harmony_umap",
              reduction.key = "UMAP_Harmony_")


so <- FindNeighbors(so, reduction = "harmony", dims = 1:40)
so <- FindClusters(so, resolution = 0.2)

so$coarse_clusters_gp34_harmony <- so$seurat_clusters
Idents(so) <- "coarse_clusters_gp34_harmony"
plot_harmony(so, "coarse_clusters_gp34_harmony")
```

```{r}
saveRDS(so, "objects/gp34_2.rds")
```

```{r}
so <- readRDS("objects/gp34_2.rds")
```


```{r markers of each subgroup}
library(presto)
mkrs <- wilcoxauc(so, "coarse_clusters_gp34_harmony")

mkrs <- filter(mkrs, logFC > 0, padj < 0.05) %>% 
  group_by(group) %>% 
  arrange(padj, desc(logFC), .by_group = TRUE)

mkrs %>% 
  write_tsv(file.path(mkrs_dir, "harmony_markers_gp34.tsv"))

mkrs %>% 
  ungroup() %>% 
  split(., .$group) %>% 
  write_markers_xlsx(.,
                   file.path(xcel_dir, "harmony_markers_gp34.xlsx"))
```

## GO terms

```{r}
library(gprofiler2)
library(ggrepel)
mkrs <- read_tsv(file.path(mkrs_dir, 
                    "harmony_markers_gp34.tsv")) %>% 
  filter( padj < 0.05) %>% 
  group_by(group) %>% 
  slice(1:500)

mkrs_split <- split(mkrs, mkrs$group) %>% 
  map(~pull(.x, feature))

go_res <- gost(mkrs_split, 
               ordered_query = TRUE,
               sources = c("GO", 
                          "KEGG", 
                          "REAC", 
                          "TF"))

go_res[["result"]] %>% 
  mutate(parents = unlist(map(parents, 
                              ~str_c(.x, collapse = ",")))) %>% 
  write_tsv(file.path(tbls_dir, "goterms_gp34_clusters.tsv"))

go_res[["result"]] %>% 
  mutate(parents = unlist(map(parents, 
                              ~str_c(.x, collapse = ","))),
         source = str_replace_all(source, 
                                  "[[:punct:]]", " ")) %>% 
  select(-significant) %>% 
  split(., .$query) %>% 
  openxlsx::write.xlsx(., file.path(tbls_dir, "goterms_gp34_cluster.xlsx"))
```

## Heatmap of gene expression

```{r}
topx <- read_tsv(file.path(mkrs_dir, 
                    "harmony_markers_gp34.tsv")) %>% 
  filter(padj < 0.05) %>% 
   mutate(group = as.factor(as.integer(group))) %>% 
  group_by(group) %>% 
  arrange(padj, desc(logFC), .by_group = TRUE) %>% 
  slice(1:10)

so <- ScaleData(so, features = rownames(so))

tmp_so <- subset(so, 
                 subset = coarse_clusters_gp34_harmony %in% as.character(c(0:5)))

tmp_so$coarse_clusters_gp34_harmony <- factor(tmp_so$coarse_clusters_gp34_harmony)

p <- DoHeatmap(tmp_so, 
               group.colors = discrete_palette_default,
               features = unique(topx$feature),
               group.by = "coarse_clusters_gp34_harmony",
               angle = 0, 
               raster = FALSE, 
               draw.lines = TRUE)

p <- p +
  scale_fill_gradientn(colours = viridis::viridis(256),
                       name = "Expression\nZ-scores")

save_plot(file.path(fig_dir, "heatmap_markers_gp34.pdf"), 
          p,
          base_width = 10,
          base_height = 12)


```

## Heatmap by TF activity

```{r}
so_tf_markers <- read_tsv("tables/gp34_tf_activities.tsv")

topx <- so_tf_markers %>% 
  filter(cluster %in% as.character(c(0:5))) %>% 
  group_by(cluster) %>% 
  arrange(p_val_adj, .by_group = TRUE) %>% 
  slice(1:5)

DefaultAssay(tmp_so) <- "TF"
tmp_so <- ScaleData(tmp_so)

p <- DoHeatmap(tmp_so, 
               group.colors = discrete_palette_default,
        features = unique(topx$gene),
        group.by = "coarse_clusters_gp34_harmony",
        angle = 0, 
        raster = FALSE, 
        draw.lines = TRUE)

p <- p +
  scale_fill_gradientn(colours = viridis::viridis(256),
                       name = "Regulon AUC\nZ-scores")

save_plot(file.path(fig_dir, "heatmap_tfs_gp34.pdf"), 
          p,
          base_width = 10,
          base_height = 6)

rm(tmp_so)
DefaultAssay(so) <- "RNA"
```

## Add in Northcutt paper metamodule scores

```{r}
supp_tbl_2 <- "https://static-content.springer.com/esm/art%3A10.1038%2Fs41586-019-1434-6/MediaObjects/41586_2019_1434_MOESM3_ESM.xlsx"

if(!file.exists("../docs/northcutt_sup_tbl_2.xlsx")){
  download.file(supp_tbl_2, "../docs/northcutt_sup_tbl_2.xlsx")
}

sup_tbl <- read_excel("../docs/northcutt_sup_tbl_2.xlsx", 
           sheet = 2,
           skip = 3) 

sup_tbl[, 1] <- NULL

# extract out each metaprogram following their approach of excluding ribosomal
# and protein genes and non-coding RNAs

mps <- map(seq(1, ncol(sup_tbl), 2), 
    ~c(.x, .x + 1)) %>% 
  map(., ~sup_tbl[, .x] 
      %>% na.omit()) 

mp_names <- map_chr(mps, ~colnames(.x)[2])
mps <- map(mps, ~.x[, 2, drop = TRUE]) 
names(mps) <- mp_names

gp34_mps <- mps[str_subset(names(mps), "Group")]

mp_ids <- c(
  "Group 3/4-A-cell_cycle",
  "Group 3/4-B-Translation_Myc",
  "Group 3/4-C-Differentiation"
)

names(gp34_mps) <- mp_ids

for (i in seq_along(gp34_mps)){
  so <- AddModuleScore(so, 
                       features = list(c(gp34_mps[[i]])),
                       ctrl = 50,
                       name = names(gp34_mps)[i],
                       seed = 42)
}

new_ids <- str_c(make.names(mp_ids), "1") 

new_id_idx <- match(new_ids, colnames(so@meta.data))
colnames(so@meta.data)[new_id_idx] <- mp_ids
```


```{r}
so <- CellCycleScoring(so,
                       g2m.features = cc.genes$g2m.genes,
                       s.features = cc.genes$s.genes)
```


## Save object and write out matrix for other tools

```{r}
saveRDS(so, "objects/gp34_2.rds")
```

```{r}
so <- readRDS("objects/gp34.rds")

#full count matrix
GetAssayData(so, "counts") %>% 
  as.matrix(.) %>% 
  as.data.frame() %>% 
  rownames_to_column("gene") %>% 
  write_tsv("tables/gp34_count_matrix.tsv.gz")
  
# expression matrix filtered at least 25 cells express the gene
expr_mat <- GetAssayData(so, "data")
expr_mat <- expr_mat[Matrix::rowSums(expr_mat > 0) >= 25, ]

expr_mat %>% 
  as.matrix(.) %>% 
  as.data.frame() %>% 
  rownames_to_column("gene") %>% 
  write_tsv("tables/gp34_expr_matrix.tsv.gz")
  
```

save loom file for paga

```{r}

# drop graphs see https://github.com/theislab/scanpy/issues/598
tmp_obj <- readRDS("objects/gp34.rds")
tmp_obj@graphs <- list()
tmp_obj@meta.data <- tmp_obj@meta.data[, c("coarse_clusters_gp34_harmony",
                                           "orig.ident",
                                           "subtype",
                                           "subgroup",
                                           "UPN")]

lfile <- as.loom(tmp_obj,filename = file.path("objects", "grp34.loom"),
        verbose = TRUE,
        overwrite = TRUE)
lfile$close_all()

rm(tmp_obj)
```

## Get top PC genes

```{r}
to_plot <- c("UPN", "coarse_cell_type", "subgroup")

dims_to_plot <- list(
  c(1,2),
  c(2,3),
  c(3,4),
  c(4,5))
  
plts <- list()
for ( i in seq_along(dims_to_plot)){
  plts[[i]] <- map(to_plot, ~plot_pca(so, .x, dims =dims_to_plot[[i]])) %>%
  plot_grid(plotlist = ., nrow = 1, ncol = length(to_plot),
            rel_widths = c(1.75, 1.5, 1))
}

names(plts) <- str_c("pc_", 
                     map_chr(dims_to_plot, ~str_c(.x, collapse = "_")),
                     ".pdf")
pc_dir <- file.path(fig_dir, "pcs_grp34")
dir.create(pc_dir)
iwalk(plts,
  ~save_plot(file.path(pc_dir, .y), .x, nrow = 1, ncol = 3))


pc_loadings <- so@reductions$pca@feature.loadings[, 1:20] %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column("gene") %>% 
  arrange(desc(PC_1))


pc_tbl_dir <- file.path(tbls_dir, "pcs", "grp34")
dir.create(pc_tbl_dir, recursive = TRUE)
write_tsv(pc_loadings, file.path(pc_tbl_dir,
                                 "pc_loadings.tsv.gz"))

tibble(pc = str_c("PC", 1:length(so@reductions$pca@stdev)), 
                  stdev = so@reductions$pca@stdev) %>% 
  write_tsv(file.path(pc_tbl_dir, "pc_variance_explained.tsv"))
```



## Average Expression per cluster

```{r}
plot_feature(so, 
             "coarse_clusters_gp34_harmony",
             embedding = "umap_harmony")

avg_tbl_dir <- file.path(tbls_dir, "avg_expr", "grp34")
dir.create(avg_tbl_dir, recursive = TRUE)

write_avg_expr(so,  
               "coarse_clusters_gp34_harmony",
               file.path(avg_tbl_dir, 
                         "avgexpr.tsv.gz"))
```

## Fix sample 1202 label

UPN 1202 is now classified as GP3 not GP3/4 based on methlyation data. 


```{r}
sg_levels <- levels(so$subgroup)
upn_levels <- levels(so$UPN)

so@meta.data <- so@meta.data %>% 
  rownames_to_column("cell") %>% 
  mutate(
    subgroup = as.character(subgroup),
    subgroup = ifelse(UPN == "1202",
                           "GP3",
                           subgroup),
    subgroup = factor(subgroup, levels = sg_levels)) %>% 
  column_to_rownames("cell")
          
plot_umap(so, "subgroup")
plot_umap(so, "UPN")

saveRDS(so, "objects/gp34.rds")
```


## CytoTRACE

```{r}
source("../R/CytoTRACE.R")
#subset to genes expressed in at least 25 cells
to_keep <- Matrix::rowSums(so@assays$RNA@counts > 0) > 25

a <- CytoTRACE(as.matrix(so@assays$RNA@counts[to_keep, ]),
               ncores = 4)

res <- tibble(
  cell = names(a$GCS),
  GCS = a$GCS,
  CytoTRACE = a$CytoTRACE) %>% 
  left_join(get_metadata(so), ., by = "cell") %>% 
  select(cell, GCS, CytoTRACE) %>% 
  mutate(GCS = ifelse(is.na(GCS), 0, GCS),
         CytoTRACE = ifelse(is.na(CytoTRACE), -1, CytoTRACE)) %>% 
  column_to_rownames("cell")

so <- AddMetaData(so, res)

plot_feature(so, 
             "CytoTRACE",
             embedding = "umap_harmony",
             sorted = FALSE)

plot_feature(so, 
             "GCS",
             embedding = "umap_harmony",
             sorted = FALSE)

plot_feature(so, 
             "nFeature_RNA",
             embedding = "umap_harmony",
             sorted = FALSE)
```

## Group3 only analysis

UPN 1202 is now classified as GP3 not GP3/4 based on methlyation data. 


```{r}
so_gp3 <- subset(so, subset = subgroup == "GP3")
so_gp4 <- subset(so, subset = subgroup == "GP4")
```


```{r}
so <- FindVariableFeatures(
  so_gp3,
  selection.method = "vst",
  nfeatures = 3000,
  verbose = FALSE
)

so <- ScaleData(so, features = VariableFeatures(so), 
                  verbose = TRUE)

so <- RunPCA(so, 
             features = VariableFeatures(so),
             npcs = 50, 
             verbose = FALSE)

ElbowPlot(so, ndims = 50)

# make graphs and use graphs for UMAP
so <- FindNeighbors(so, 
                      reduction = "pca", 
                      dims = 1:50, 
                      k.param = 50L)


so <- RunUMAP(so, 
              dims = 1:50, 
              n.neighbors = 30L,
              min.dist = 0.4)

so <- FindClusters(so, 
                   resolution = c(0.1, 0.3, 0.5))

so$coarse_clusters_gp3 <- so$RNA_snn_res.0.1

Idents(so) <- "coarse_clusters_gp3"

plot_umap(so, "coarse_clusters_gp3")
plot_umap(so, "UPN")
plot_umap(so, "subtype")
```

## Harmonize

```{r}
library(harmony)
so <- RunHarmony(so,
                 "UPN", 
                 theta = 1.5,
                 plot_convergence = TRUE)

so <- RunUMAP(so, 
              reduction = "harmony", 
              dims = 1:50, 
              min.dist = 0.3,
              n.neighbors = 30L,
              reduction.name = "umap_harmony",
              reduction.key = "UMAP_Harmony_")

plot_feature(so, "UPN", embedding = "umap_harmony")
plot_umap(so, "UPN")
plot_umap(so, "subtype")

so <- FindNeighbors(so, reduction = "harmony", dims = 1:40)
so <- FindClusters(so, resolution = 0.2)


so$coarse_clusters_gp3_harmony <- so$seurat_clusters
Idents(so) <- "coarse_clusters_gp3_harmony"
plot_feature(so, "coarse_clusters_gp3_harmony", embedding = "umap_harmony")
plot_feature(so, "NRL", embedding = "umap_harmony")
```


```{r}
library(presto)
library(scbp)
mkrs <- wilcoxauc(so, "coarse_clusters_gp3_harmony")

filter(mkrs, 
       logFC > 0, padj < 0.05) %>% 
  group_by(group) %>% 
  arrange(padj, desc(logFC)) %>% 
write_tsv(., 
          file.path(mkrs_dir, "harmony_markers_gp3_only.tsv"))

filter(mkrs, 
       logFC > 0, padj < 0.05) %>% 
  group_by(group) %>% 
  arrange(padj, desc(logFC)) %>% 
  ungroup() %>% 
  split(., .$group) %>% 
scbp::write_markers_xlsx(.,
                   file.path(xcel_dir, "harmony_markers_gp3_only.xlsx"))
```

```{r}
saveRDS(so, "objects/gp3_only.rds")
```

```{r, eval = FALSE}
so <- readRDS("objects/gp3_only.rds")
```
## GP4 only

```{r}
so <- FindVariableFeatures(
  so_gp4,
  selection.method = "vst",
  nfeatures = 3000,
  verbose = FALSE
)

so <- ScaleData(so, features = VariableFeatures(so), 
                  vars.to.regress = c("percent.mt", "nCount_RNA"), 
                  verbose = TRUE)

so <- RunPCA(so, 
             features = VariableFeatures(so),
             npcs = 50, 
             verbose = FALSE)

ElbowPlot(so, ndims = 50)

# make graphs and use graphs for UMAP
so <- FindNeighbors(so, 
                      reduction = "pca", 
                      dims = 1:50, 
                      k.param = 50L)


so <- RunUMAP(so, 
              dims = 1:50, 
              n.neighbors = 30L,
              min.dist = 0.4)

so <- FindClusters(so, 
                   resolution = c(0.1, 0.3, 0.5))

so$coarse_clusters_gp4 <- so$RNA_snn_res.0.1

Idents(so) <- "coarse_clusters_gp4"

plot_umap(so, "coarse_clusters_gp4")
plot_umap(so, "UPN")
plot_umap(so, "subtype")
```

## Harmonize

```{r}
so <- RunHarmony(so,
                 "UPN", 
                 theta = 1.5,
                 plot_convergence = TRUE)

so <- RunUMAP(so, 
              reduction = "harmony", 
              dims = 1:50, 
              min.dist = 0.3,
              n.neighbors = 30L,
              reduction.name = "umap_harmony",
              reduction.key = "UMAP_Harmony_")

plot_feature(so, "UPN", embedding = "umap_harmony")
plot_umap(so, "UPN")
plot_umap(so, "subtype")

so <- FindNeighbors(so, reduction = "harmony", dims = 1:40)
so <- FindClusters(so, resolution = 0.2)


so$coarse_clusters_gp4_harmony <- so$seurat_clusters
Idents(so) <- "coarse_clusters_gp4_harmony"
plot_feature(so, "coarse_clusters_gp4_harmony", embedding = "umap_harmony")
```

```{r}
library(presto)
library(scbp)
mkrs <- wilcoxauc(so, "coarse_clusters_gp4_harmony")

filter(mkrs, 
       logFC > 0, padj < 0.05) %>% 
  group_by(group) %>% 
  arrange(padj, desc(logFC)) %>% 
  ungroup() %>% 
write_tsv(., 
          file.path(mkrs_dir, "harmony_markers_gp4_only.tsv"))

filter(mkrs, 
       logFC > 0, padj < 0.05) %>% 
  group_by(group) %>% 
  arrange(padj, desc(logFC)) %>% 
  ungroup() %>% 
  split(., .$group) %>% 
scbp::write_markers_xlsx(.,
                   file.path(xcel_dir, "harmony_markers_gp4_only.xlsx"))
```


```{r}
saveRDS(so, "objects/gp4_only.rds")
```

```{r, eval = FALSE}
so <- readRDS("objects/gp4_only.rds")
```