---
title: "11_CNV"
author: "Kent Riemondy RBI"
date: "`r Sys.Date()`"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "html") })
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: cosmo
    highlight: tango
    fig_caption: true
    code_folding: hide
    df_print: paged
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libs}
source(here::here("/R/utils.R"))
library(ComplexHeatmap)
library(presto)
library(openxlsx)
library(gprofiler2)
library(data.table)

rmd_dir <- "cnvs"

fig_dir <- file.path(fig_dir, rmd_dir)

walk(fig_dir,
     dir.create, 
     showWarnings = FALSE)
```

## CNVS from methylation data


```{r}
# from methylation data

methyl_cnvs <- tibble::tribble(
  ~'UPN', ~'methyl_nCNV',
    "925",   6L,
    "945",   7L,
   "1028",   3L,
   "1130",   6L,
   "1167",   3L,
   "1355",   7L,
   "1433",   8L,
    "753",   1L,
    "934",   1L,
    "966",   4L,
    "996",  17L,
   "1066",   2L,
   "1070",   4L,
   "1125",   9L,
   "1155",   7L,
   "1177",   2L,
   "1195",   4L,
   "1238",   9L,
  "966_2",   5L,
    "801",   9L,
    "831",   2L,
    "877",   2L,
    "898",   3L,
   "1224",   0L,
   "1235",   3L,
   "1325",   6L,
   "1397",  14L,
   "1416",   1L,
   "1128",   1L
  )

```

## Strategy

Tumors cells tend to cluster into largely patient specific clusters. However normal cells (i.e. immune and astrocytes) tend to cluster with cells from multiple patients. This argues that the heterogeneity present in each patient is not simply an artifact from experimental processing, but insteads reflects intrinsic heterogeneity. This hetergeneity could be driven by enviromental factors (tumor location, aggressivness, etc), or tumor genetics. Each patients CNV profile is hetergeneous, suggesting that CNV could driven expression variation, providing an explanation of the hetergeneous expression profiles.

To assess this possibility, inferCNV was run to identify CNVs. The genes in these CNV regions will be compared to marker genes of each patients tumor. Need to ensure that markers that are down regulated are enriched in CNV regions that are lost and vice versa.

Here is the coding for the inferCNV hmm results:
1: complete loss
2: loss of one copy
3: neutral
4: addition of one copy
5: addition of two copies
6: essentially a placeholder for >2x copies but modeled as 3x.


```{r}
mkrs <- read_tsv(file.path(mkrs_dir, 
                           "preprocess",
                           "upn_markers_neoplastic_pos_and_neg.tsv"),
                 col_types = c("ccnnnnnnnn")) %>% 
  filter(padj < 0.01,
         pct_in > 0.10,
         abs(logFC) > log(1.5))

subgroups <- c("shh","gp3", "gp4")
names(subgroups) <- c("shh", "gp3", "gp4")
cnv_reports <- map_dfr(subgroups,
                   ~read_tsv(file.path("inferCNV_v2",
                                       .x, 
                                      .x,
                   "HMM_CNV_predictions.HMMi6.hmm_mode-samples.Pnorm_0.5.pred_cnv_genes.dat")),
                   .id = "subgroup") 

cnv_reports <- mutate(cnv_reports, 
                      upn = str_split(cell_group_name,
                                      "\\.", 
                                      simplify = T)[, 1] %>%
                        str_remove("malignant_"),
                      change = case_when(
                        state < 3 ~ "loss",
                        state > 3 ~ "gain",
                        TRUE ~ "neutral"
                                      )) 

# get list of gnes up and down per UPN
mkr_gene_list <- split(mkrs, mkrs$group) %>% 
  map(~list(loss = filter(.x, logFC < 0) %>% pull(feature), 
            gain = filter(.x, logFC > 0) %>% pull(feature) ))


cnv_gene_list <- split(cnv_reports, cnv_reports$upn) %>%
   map(~list(loss = filter(.x, change == "loss") %>% pull(gene), 
            gain = filter(.x, change == "gain") %>% pull(gene)))

upns <- intersect(names(mkr_gene_list), names(cnv_gene_list))
```


```{r}

calc_enrichment <- function(sample_success,
                            pop_success,
                            pop_size,
                            sample_size){
  x <- sample_success - 1
  m <- pop_success 
  n <- pop_size
  k <- sample_size
  1-phyper(x, m, n, k)
}

#get some idea of a good background number

so <- qread("objects/preprocess/all_neoplastic.qs")

n_features_per_patient <- map_int(upns, 
                             function(upn){
                               tmp <- subset(so, cells = rownames(so@meta.data)[so$orig.ident == upn])
                               sum(Matrix::rowSums(GetAssayData(tmp, "counts")) > 0)
                             })

gene_pop_size <- median(n_features_per_patient)

res <- map_dfr(upns, 
    function(upn){
      x <- mkr_gene_list[[upn]]$loss
      y <- cnv_gene_list[[upn]]$loss
      pop_success_loss <- length(y)
      sample_success_loss <- length(intersect(x, y))
      sample_size_loss <- length(x)
      
      x <- mkr_gene_list[[upn]]$gain
      y <- cnv_gene_list[[upn]]$gain
      pop_success_gain <- length(y)
      sample_success_gain <- length(intersect(x, y))
      sample_size_gain <- length(x)
      
      sample_success <- sample_success_gain + sample_success_loss
      pop_success <- pop_success_gain + pop_success_loss
      sample_size <- sample_size_gain + sample_size_loss
      
      pop_size <- gene_pop_size
      tibble(upn = upn,
             n_shared = sample_success,
             n_cnv_genes = pop_success,
             n_mrks = sample_size,
             pval = 1- phyper(sample_success - 1,
                             pop_success, 
                             pop_size, 
                             sample_size))
    }) 

res <- res %>% 
  mutate(prop_markers_in_cnv = n_shared / n_mrks,
         prop_genes_in_cnv_in_genome = n_cnv_genes / gene_pop_size,
         log2_effect_size = log2(prop_markers_in_cnv / prop_genes_in_cnv_in_genome))

res
```


```{r}
mdata <- so@meta.data %>% 
  as_tibble() %>% 
  select(UPN, subgroup, tx_nCNV) %>%
  mutate(subgroup = ifelse(UPN == "1167", "GP3", subgroup)) %>% 
  unique() 

plt_dat <- res %>% 
  left_join(mdata, by = c("upn" = "UPN")) %>% 
  select(upn,
         tx_nCNV,
         log2_effect_size,
         pvalue = pval,
         subgroup,
         `Marker genes within a CNV` = prop_markers_in_cnv, 
         `All genes within a CNV` = prop_genes_in_cnv_in_genome) %>% 
  pivot_longer(cols = -c(upn, subgroup, pvalue, log2_effect_size, tx_nCNV))
  # left_join(methyl_cnvs, by = c("upn" = "UPN"))

ncnv_plt_dat <- select(plt_dat, upn, subgroup, tx_nCNV) %>% 
  unique() %>% 
  group_by(subgroup) %>% 
  arrange(tx_nCNV, .by_group = T) %>% 
  mutate(upn = factor(upn, levels = upn))

plt_dat <- mutate(plt_dat, upn = factor(upn, levels = ncnv_plt_dat$upn))

pval_plt_dat <- plt_dat %>% 
  select(upn, pvalue, subgroup, value) %>% 
  unique() %>% 
  group_by(upn) %>% 
  mutate(sig = ifelse(pvalue < 1e-100,
                      "***",
                      ifelse(pvalue < 1e-10,
                             "**",
                             ifelse(pvalue < 0.01,
                                    "*",
                                    ""))),
         value = max(value) * 1.05) 

p <- ggplot(plt_dat, aes(upn, value)) +
  geom_col(aes(fill = name), position = position_dodge2()) +
  geom_text(data = pval_plt_dat, 
            aes(label = sig)) + 
  labs(x = "",
       y = "Proportion") +
  facet_grid(subgroup ~ ., space = "free", scales = "free") + 
  scale_fill_manual(name = "",
                    values = palette_OkabeIto) + 
 coord_flip() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        legend.position = "top",
        strip.text = element_text(size = 10))

p


p2 <- ggplot(ncnv_plt_dat, aes(upn, tx_nCNV)) +
  geom_col(aes(fill = subgroup), position = position_dodge2()) +
  labs(x = "Tumor Id",
       y = "# of CNVs") +
  facet_grid(subgroup ~ ., space = "free", scales = "free") + 
  scale_fill_manual(name = "",
                    values = palette_OkabeIto) + 
 coord_flip() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        legend.position = "top",
        strip.text = element_blank())

p <- plot_grid(p2, p)
save_plot("figs/cnvs/cnv_marker_enrichment.pdf", p, 
          base_asp = 1, base_height = 5.5,
          nrow = 1,
          ncol = 2)


```
For most tumors there is a signficant overlap between marker genes and CNV regions. 1202 and 1224 have non-significant overlap. 


# Examine subclones

For some tumors it appears that there are some subclones based on CNV. For this anslysis I will split each tumor into 2 subclones based on HClustering on the CNV results. Then gene expression between theses splits will be evaluated to see if there are DE genes. 

First, regenerate heatmap from inferCNV to save as hi-resolution.


```{r}
library(ape)
library(dendextend)
library(scbp)
library(RColorBrewer)
library(ComplexHeatmap)

subgroups <- c(
  "gp3",
  "gp4",
  "shh",
  "all"
)

trees <- map(subgroups,
             ~read.tree(
               file.path("inferCNV_v2",
             .x, .x,  "infercnv.observations_dendrogram.txt"))) 

ref_dat <- map(subgroups,
             ~read.table(
               file.path("inferCNV_v2",
             .x, .x,  "infercnv.references.txt"), 
             sep = " ") %>% 
               as.matrix() %>% 
               t()) 

ob_dat <- map(subgroups,
             ~read.table(
               file.path("inferCNV_v2",
             .x, .x,  "infercnv.observations.txt"), 
             sep = " ") %>% 
               as.matrix() %>% 
               t()) 

gene_df <- read_tsv("inferCNV_v2/gene_ordering_file.txt",
                    col_names = c("gene", "chrom", "start", "stop"))

# merge per tumor dendograms for each subgroup
ds <- map(trees, 
    ~map(.x, as.dendrogram)) %>% 
  map(., 
      ~Reduce(function(x, y) {merge(x, y)}, .x))

# save heatmap inputs as rds files 

dir.create(file.path("objects", "cnv"))
for(i in seq_along(subgroups)){
  saveRDS(list(
    ob = ob_dat[[i]],
    ref = ref_dat[[i]],
    genes = gene_df,
    dend =  ds[[i]]
  ),
  file.path("objects", "cnv", paste0(subgroups[i], "_hmap.rds"))
  )
}

```



```{r}
plot_cnv_heatmap <- function(ob_dat, 
                             ref_dat,
                             gene_df, 
                             dend, 
                             grp = "SHH", 
                             grp_col = palette_OkabeIto[3]){
  
  col_fun <- circlize::colorRamp2(c(min(ob_dat), median(ob_dat), max(ob_dat)),
                                  c("blue", "#FFFFFF", "red"))
  
  chr_annot <- left_join(tibble(gene = colnames(ob_dat)),
                         gene_df) %>%
    mutate(chrom = str_remove(chrom, "chr"), 
           chrom = factor(chrom, levels = unique(chrom))) %>% 
    column_to_rownames("gene") %>% 
    select(-start, -stop) 

  
  ob_annot <- data.frame(
    row.names = rownames(ob_dat),
    #subgroup = grp,
    tumor = str_remove(rownames(ob_dat), "^X") %>% 
      str_split(., "_", simplify = TRUE) %>% .[ ,1],
    stringsAsFactors = FALSE
  )
  
  tumor_order <- sort(unique(as.character(ob_annot$tumor)))
  
  subgroup_cols <-  palette_OkabeIto[1]
  names(subgroup_cols) <- grp
  
  ha_ob <- rowAnnotation(df = ob_annot, 
                         col = list(
                           subgroup = structure(
                             grp_col,
                             names = grp
                           ),
                           tumor = structure(
                             discrete_palette_default[1:length(tumor_order)],
                             names = tumor_order)
                         ),
                         annotation_legend_param = list(by_row = TRUE, 
                                                      nrow = 2))
  
  ob_hmap <- Heatmap(ob_dat, 
                     heatmap_legend_param = list(title = ""),
                     col = col_fun,
                     cluster_columns = FALSE, 
                     show_row_names =  FALSE,
                     show_column_names = FALSE, 
                     cluster_rows = dend,
                     left_annotation = ha_ob,
                     use_raster = TRUE,
                     column_split = chr_annot$chrom,
                     column_gap = unit(0, "mm"),
                     column_title_gp = gpar(fontsize = 2),
                     border = TRUE,
                     row_title = "Tumor",
                     raster_device = "png",
                     raster_quality = 3)
  
  ref_hmap <- Heatmap(ref_dat, 
                      col = col_fun,
                      cluster_columns = FALSE, 
                      show_row_names =  FALSE,
                      show_column_names = FALSE, 
                      cluster_rows = FALSE,
                      use_raster = TRUE,
                      column_split = chr_annot$chrom,
                      show_heatmap_legend = FALSE,
                      border = TRUE,
                      column_gap = unit(0, "mm"),
                      row_title = "Normal",
                      raster_device = "png",
                      raster_quality = 3)
  
  ht_list <- ref_hmap %v% ob_hmap
  
  dynamic_resize <- 1
  nobs <- 100
  pdf(str_c("hmap_", grp, ".pdf"),
      useDingbats=FALSE,
      width=11,
      height = 9 + dynamic_resize * 3.6 * (nobs - 200)/200 )
  draw(ht_list, heatmap_legend_side = "right", annotation_legend_side = "bottom")
  dev.off()
  
  list(ht_list, ob_hmap)
}

for(i in seq_along(subgroups)){
  x <- readRDS(file.path("objects", "cnv", paste0(subgroups[i], "_hmap.rds")))
   
  plts <-  plot_cnv_heatmap(x$ob,
                            x$ref, 
                            x$genes,
                            x$dend, 
                            grp = subgroups[i],
                            grp_col = palette_OkabeIto[i])
  
  saveRDS(plts, file.path("objects", "cnv", paste0(subgroups[i], "_hmap_plots.rds")))
  rm(x,plts)
  gc()
}

```

make heatmap with all
```{r}
x <- readRDS(file.path("objects", "cnv", "all_hmap.rds"))
ob_dat <- x$ob
ref_dat <- x$ref
gene_df <- x$genes
dend <- x$dend
trees <- read.tree(
               file.path("inferCNV_v2",
             "all", "all",  "infercnv.observations_dendrogram.txt")) %>% 
  map(., as.dendrogram)

names(trees) <- map_chr(trees, 
                        ~str_split(labels(.x), "_", simplify = T) %>% 
                          .[, 1] %>% 
                          ifelse(. == "943", "934", .) %>% 
                          ifelse(. == "966-2", "966", .) %>% 
                          unique() 
                          )
             
grps <- c("GP3", "GP4", "SHH", "WNT")
grp_cols <- palette_OkabeIto[1:4]


col_fun <- circlize::colorRamp2(c(min(ob_dat), median(ob_dat), max(ob_dat)),
                                  c("blue", "#FFFFFF", "red"))
  
mdata <- read_excel(file.path(doc_dir, 
                                   "MED scRNAseq batch and annotations 5.8.19.xlsx")) %>% 
  filter(subgroup != "LGG") %>% 
  select(-subgroup)

mdata_updated <- read_excel(file.path(doc_dir, 
                                   "MED scRNAseq batch and annotations no TBD 5.28.19.xlsx")) %>% 
  rename(subgroup = `subgroup by transcriptomics` , 
         date = ...7,
         fq_id = ...8 ) %>% 
  filter(subgroup != "LGG") %>% 
  select(UPN, subgroup)

mdata <- left_join(mdata_updated, mdata, by = "UPN")

ob_annot <- tibble(
  ids = rownames(ob_dat),
  tumor = str_remove(rownames(ob_dat), "^X") %>% 
    str_split(., "_", simplify = TRUE) %>% .[ ,1]
) %>% 
  left_join(mdata, by = c("tumor" = "UPN")) %>% 
  select(ids, tumor, subgroup) %>%
  mutate(subgroup = ifelse(subgroup == "GP3/4", "GP3", subgroup),
         subgroup = ifelse(tumor == "966.2", "GP4", subgroup),
         tumor = ifelse(tumor == "966.2", "966", tumor),
         tumor = ifelse(tumor == "943", "934", tumor)) %>% 
  group_by(tumor) %>%  
  column_to_rownames("ids") %>% 
  as.data.frame(stringsAsFactors = FALSE) 

# 
 # ob_dat <- ob_dat[rownames(ob_annot), 1:100]
 # ref_dat <- ref_dat[1:10, 1:100]

chr_annot <- left_join(tibble(gene = colnames(ob_dat)),
                       gene_df) %>%
  mutate(chrom = str_remove(chrom, "chr"), 
         chrom = factor(chrom, levels = unique(chrom))) %>% 
  column_to_rownames("gene") %>% 
  select(-start, -stop) 



tumor_order <- sort(unique(as.character(ob_annot$tumor)))

subgroup_cols <-  grp_cols
names(subgroup_cols) <- grps

ha_ob <- rowAnnotation(df = ob_annot, 
                       col = list(
                         subgroup = structure(
                           grp_cols,
                           names = grps
                         ),
                         tumor = structure(
                           discrete_palette_default[1:length(tumor_order)],
                           names = tumor_order)
                       ),
                       annotation_legend_param = list(by_row = TRUE, 
                                                      nrow = 2))


dend <- map(split(ob_annot, ob_annot$subgroup),
            ~{
     d <- as.dendrogram(hclust(dist(ob_dat[rownames(.x), ]))) %>% 
     Reduce(function(x, y) {merge(x, y)}, .)

ob_hmap <- Heatmap(ob_dat, 
                   heatmap_legend_param = list(title = ""),
                   col = col_fun,
                   cluster_columns = FALSE, 
                   show_row_names =  FALSE,
                   show_column_names = FALSE, 
                   cluster_rows = FALSE,
                   split = ob_annot$subgroup,
                   left_annotation = ha_ob,
                   column_split = chr_annot$chrom,
                   column_gap = unit(0, "mm"),
                   column_title_gp = gpar(fontsize = 2),
                   cluster_row_slices = FALSE,
                   border = TRUE,
                   row_title = "Tumor",
                   use_raster = TRUE,
                   raster_device = "png",
                   raster_quality = 3)

ref_hmap <- Heatmap(ref_dat, 
                    col = col_fun,
                    cluster_columns = FALSE, 
                    show_row_names =  FALSE,
                    show_column_names = FALSE, 
                    cluster_rows = FALSE,
                    use_raster = TRUE,
                    column_split = chr_annot$chrom,
                    show_heatmap_legend = FALSE,
                    border = TRUE,
                    column_gap = unit(0, "mm"),
                    row_title = "Normal",
                    raster_device = "png",
                    raster_quality = 3)

ht_list <- ref_hmap %v% ob_hmap

dynamic_resize <- 1
nobs <- nrow(ob_dat)
nobs <- 100
pdf(str_c("hmap_all_sub.pdf"),
    useDingbats=FALSE,
    width=11,
    height = 9 + dynamic_resize * 3.6 * (nobs - 200)/200 )
draw(ht_list, heatmap_legend_side = "right", annotation_legend_side = "bottom")
dev.off()

list(ht_list, ob_hmap)

```


make heatmap with all with subgroup splits
```{r}
x <- readRDS(file.path("objects", "cnv", "all_hmap.rds"))
ob_dat <- x$ob
ref_dat <- x$ref
gene_df <- x$genes
dend <- x$dend
trees <- read.tree(
               file.path("inferCNV_v2",
             "all", "all",  "infercnv.observations_dendrogram.txt")) %>% 
  map(., as.dendrogram)

names(trees) <- map_chr(trees, 
                        ~str_split(labels(.x), "_", simplify = T) %>% 
                          .[, 1] %>% 
                          ifelse(. == "943", "934", .) %>% 
                          ifelse(. == "966-2", "966", .) %>% 
                          unique() 
                          )
             
grps <- c("GP3", "GP4", "SHH", "WNT")
grp_cols <- palette_OkabeIto[1:4]


col_fun <- circlize::colorRamp2(c(min(ob_dat), median(ob_dat), max(ob_dat)),
                                  c("blue", "#FFFFFF", "red"))
  
mdata <- read_excel(file.path(doc_dir, 
                                   "MED scRNAseq batch and annotations 5.8.19.xlsx")) %>% 
  filter(subgroup != "LGG") %>% 
  select(-subgroup)

mdata_updated <- read_excel(file.path(doc_dir, 
                                   "MED scRNAseq batch and annotations no TBD 5.28.19.xlsx")) %>% 
  rename(subgroup = `subgroup by transcriptomics` , 
         date = ...7,
         fq_id = ...8 ) %>% 
  filter(subgroup != "LGG") %>% 
  select(UPN, subgroup)

mdata <- left_join(mdata_updated, mdata, by = "UPN")

ob_annot <- tibble(
  ids = rownames(ob_dat),
  tumor = str_remove(rownames(ob_dat), "^X") %>% 
    str_split(., "_", simplify = TRUE) %>% .[ ,1]
) %>% 
  left_join(mdata, by = c("tumor" = "UPN")) %>% 
  select(ids, tumor, subgroup) %>%
  mutate(subgroup = ifelse(subgroup == "GP3/4", "GP3", subgroup),
         subgroup = ifelse(tumor == "966.2", "GP4", subgroup),
         tumor = ifelse(tumor == "966.2", "966", tumor),
         tumor = ifelse(tumor == "943", "934", tumor)) %>% 
  group_by(tumor) %>%  
  column_to_rownames("ids") %>% 
  as.data.frame(stringsAsFactors = FALSE) 



subgroup_split <- split(ob_annot, ob_annot$subgroup)

hmaps <- map(subgroup_split,
             ~{
               ids <- unique(.x$tumor)
               
               d <- trees[ids] %>% 
                 Reduce(function(x, y) {merge(x, y)}, .)
               
               ob_to_plot <- ob_dat[rownames(.x), ]
               
               chr_annot <- left_join(tibble(gene = colnames(ob_to_plot)),
                                      gene_df) %>%
                 mutate(chrom = str_remove(chrom, "chr"), 
                        chrom = factor(chrom, levels = unique(chrom))) %>% 
                 column_to_rownames("gene") %>% 
                 select(-start, -stop) 
               
               ha_ob <- rowAnnotation(df = ob_annot[rownames(.x), ], 
                                      col = list(
                                        subgroup = structure(
                                          grp_cols,
                                          names = grps
                                        ),
                                        tumor = structure(
                                          discrete_palette_default[1:length(tumor_order)],
                                          names = tumor_order)
                                      ),
                                      annotation_legend_param = list(by_row = TRUE, 
                                                                     nrow = 2))
               
               ob_hmap <- Heatmap(ob_to_plot, 
                                  heatmap_legend_param = list(title = ""),
                                  col = col_fun,
                                  cluster_columns = FALSE, 
                                  show_row_names =  FALSE,
                                  show_column_names = FALSE, 
                                  cluster_rows = d,
                                  #  split = ob_annot$subgroup,
                                  left_annotation = ha_ob,
                                  column_split = chr_annot$chrom,
                                  column_gap = unit(0, "mm"),
                                  column_title_gp = gpar(fontsize = 2),
                                  cluster_row_slices = FALSE,
                                  border = TRUE,
                                  show_heatmap_legend = unique(.x$subgroup) == "WNT",
                                  row_title = unique(.x$subgroup),
                                  use_raster = TRUE,
                                  raster_device = "png",
                                  raster_quality = 3)
             })

ref_hmap <- Heatmap(ref_dat, 
                    col = col_fun,
                    cluster_columns = FALSE, 
                    show_row_names =  FALSE,
                    show_column_names = FALSE, 
                    cluster_rows = FALSE,
                    use_raster = TRUE,
                    column_split = chr_annot$chrom,
                    show_heatmap_legend = FALSE,
                    border = TRUE,
                    column_gap = unit(0, "mm"),
                    row_title = "Normal",
                    raster_device = "png",
                    raster_quality = 3)

ht_list <- ref_hmap %v% hmaps[[1]] %v% hmaps[[2]] %v% hmaps[[3]] %v% hmaps[[4]] 

dynamic_resize <- 1
nobs <- nrow(ob_dat)
nobs <- 100
pdf(str_c("hmap_all.pdf"),
    useDingbats=FALSE,
    width=11,
    height = 9 + dynamic_resize * 3.6 * (nobs - 200)/200 )
draw(ht_list, heatmap_legend_side = "right", annotation_legend_side = "bottom")
dev.off()


```




```{r}
z <- labels(ds[[1]]) %>% str_split("_", simplify = TRUE) %>% .[, 1] %>% as.numeric(as.factor(.))

a <- color_branches(ds[[1]], clusters = z) 
plot(a)

map(trees, 
    ~map(.x, ~dendextend::cutree(.x, k = 2)) %>% unlist()) %>%
  unlist() -> a

so$cnv <- a[colnames(so)]
so$cnv <- ifelse(is.na(so$cnv), -1,so$cnv)
```

```{r}

to_plot <- c(
  "cnv",
  "UPN",
  "nCount_RNA",
  "subgroup"
)

map2(to_plot, 
    c("by_feature",
      "random",
      "random",
      "random"),
    ~plot_feature(so, 
               .x, 
               pt_alpha = 0.7, 
               sorted = .y,
               embedding = "umap")) %>% 
  plot_grid(plotlist = .)
```

```{r}
plot_violin(get_metadata(so), "cnv", "nCount_RNA")

```

## Add CNV data to seurat objects/browser

```{r}

so_obj_fns <- c(
  shh = file.path("objects", "shh", "shh.qs"),
  gp3 = file.path("objects", "gp34", "gp3_only.qs"),
  gp4 = file.path("objects", "gp34", "gp4_only.qs")
)

sos <- imap(so_obj_fns,~{
  tmp_so <- qread(.x, nthreads = 4)
  tmp <- infercnv::add_to_seurat(infercnv_output_path = file.path("inferCNV_v2", .y, .y))
  hmm_pred <- read.table(file.path("inferCNV_v2", .y, .y, "map_metadata_from_infercnv.txt")) %>% 
    as.matrix()
  to_keep <- intersect(colnames(tmp_so), rownames(hmm_pred))
  missing_cells <- setdiff(colnames(tmp_so), rownames(hmm_pred))
  hmm_pred <- t(hmm_pred[to_keep, ])
  
  na_vals <- matrix(0L, 
                    nrow = nrow(hmm_pred), 
                    ncol = length(missing_cells))
  rownames(na_vals) <- rownames(hmm_pred)
  colnames(na_vals) <- missing_cells
  hmm_pred <- cbind(hmm_pred, na_vals)
  hmm_pred <- hmm_pred[, colnames(tmp_so)] 
  
  tmp_so[["CNV"]] <- CreateAssayObject(data = hmm_pred)
  
  if("methyl_nCNV" %in% colnames(tmp_so@meta.data)){
    tmp_so$methyl_nCNV <- NULL
  }
  
  tmp_so <- get_metadata(tmp_so) %>% 
    select(cell, UPN) %>% 
    left_join(methyl_cnvs, by = "UPN") %>% 
    column_to_rownames("cell") %>% 
    AddMetaData(tmp_so, .)

  tmp_so
})


cnv_call_subset <- str_subset(rownames(sos$shh@assays$CNV@data), "has-cnv")

# from infercnv github:
# The fields has_cnv/has_loss/has_dupli are 0/1 boolean answers to whether respectively any kind of cnv, a loss or a duplication is found within the given chr.
# proportion_cnv/proportion_loss/proportion_dupli is the proportion in number of genes that are part of any cnv/loss cnv/duplication cnv in the given chr.
# proportion_scaled are similar to proportion but taking into account whether genes that are duplications/losses are of a single copy or two copies as a weight.
# top_loss_n/top_dupli_n are the n loss/duplication cnvs that cover the most genes.

tmp <- c("has-cnv-chr9",
  "has-loss-chr9",
  "has-dupli-chr9",
  "proportion-cnv-chr9",
  "proportion-loss-chr9",
  "proportion-dupli-chr9",
  "proportion-scaled-cnv-chr9",
  "proportion-scaled-loss-chr9",
  "proportion-scaled-dupli-chr9")

plts <- map(sos, ~plot_umap(.x, c("methyl_nCNV", cnv_call_subset)) %>% 
              plot_grid(plotlist = ., 
                        nrow = 6, 
                        ncol = 4,
                        base_asp = 1.4))

iwalk(plts,
     ~save_plot(file.path(fig_dir,
                          str_c(.y, "_cnv_overview.pdf")),
                          .x,
                nrow = 6,
                ncol = 4,
                base_asp = 1.4))

walk2(sos, so_obj_fns, ~qsave(.x, .y))
```


also add to all neoplastic

```{r}
so <- qread( file.path("objects", "preprocess", "all_neoplastic.qs"))

so <- get_metadata(so) %>% 
    select(cell, UPN) %>% 
    left_join(methyl_cnvs, by = "UPN") %>% 
    column_to_rownames("cell") %>% 
    AddMetaData(so, .)
 
plot_umap(so, "methyl_nCNV")

subgroups <- c("shh","gp3", "gp4")
names(subgroups) <- c("shh", "gp3", "gp4")
cnv_reports <- map_dfr(subgroups,
                   ~read_tsv(file.path("inferCNV_v2",
                                       .x, 
                                      .x,
                   "HMM_CNV_predictions.HMMi6.hmm_mode-samples.Pnorm_0.5.pred_cnv_genes.dat")),
                   .id = "subgroup") %>% 
  mutate(UPN = str_remove(cell_group_name, "^malignant_") %>% 
           str_split("\\.", simplify = T) %>% .[, 1])



nCNV <- cnv_reports %>%
  filter(state != 3) %>% 
  group_by(subgroup, UPN) %>% 
  summarize(tx_nCNV = n_distinct(gene_region_name))

so$tx_nCNV <- get_metadata(so) %>% 
  select(UPN, coarse_cell_type) %>% 
   left_join(nCNV, by = "UPN") %>% 
  mutate(tx_nCNV = ifelse(coarse_cell_type == "malignant",
                          tx_nCNV,
                          0L)) %>% 
  pull(tx_nCNV)

qsave(so, file.path("objects", "preprocess", "all_neoplastic.qs"))
```



also add to all data

```{r}
so <- qread( file.path("objects", "preprocess", "so.qs"))

so$methyl_nCNV <- get_metadata(so) %>% 
  select(UPN, coarse_cell_type) %>% 
  left_join(methyl_cnvs, by = "UPN") %>% 
  mutate(methyl_nCNV = ifelse(coarse_cell_type == "malignant",
                              methyl_nCNV,
                              0L)) %>% 
  pull(methyl_nCNV)
 
p <- plot_umap(so, "methyl_nCNV", legend_title = "")
p
save_plot(file.path("figs", "all_data", "umap_by_methylation_cnvs.pdf"),
          p, 
          base_asp = 1.2)

subgroups <- c("shh","gp3", "gp4")
names(subgroups) <- c("shh", "gp3", "gp4")
cnv_reports <- map_dfr(subgroups,
                   ~read_tsv(file.path("inferCNV_v2",
                                       .x, 
                                      .x,
                   "HMM_CNV_predictions.HMMi6.hmm_mode-samples.Pnorm_0.5.pred_cnv_genes.dat")),
                   .id = "subgroup") %>% 
  mutate(UPN = str_remove(cell_group_name, "^malignant_") %>% 
           str_split("\\.", simplify = T) %>% .[, 1])


nCNV <- cnv_reports %>%
  filter(state != 3) %>% 
  group_by(subgroup, UPN) %>% 
  summarize(tx_nCNV = n_distinct(gene_region_name))

so$tx_nCNV <- get_metadata(so) %>% 
  select(UPN, coarse_cell_type) %>% 
   left_join(nCNV, by = "UPN") %>% 
  mutate(tx_nCNV = ifelse(coarse_cell_type == "malignant",
                          tx_nCNV,
                          0L)) %>% 
  pull(tx_nCNV)

p <- plot_umap(so, "tx_nCNV", legend_title = "")
p
save_plot(file.path("figs", "all_data", "umap_by_tx_cnvs.pdf"),
          p, 
          base_asp = 1.2)

```


```{r}
qsave(so, file.path("objects", "preprocess", "so.qs"))
```

# run once inferCNV finishes
      
```{r}              
# hmm_pred <- read_tsv("inferCNV_v2/gp4/gp4/HMM_CNV_predictions.HMMi6.hmm_mode-samples.Pnorm_0.5.pred_cnv_regions.dat") %>% 
#   separate(cell_group_name, c("id1", "id2"), sep = "\\.") %>% 
#   mutate(UPN = str_remove(id1, "malignant_")) %>% 
#   select(-id1, -id2) %>% 
#   group_by(UPN) %>% 
#   summarize(tx_nCNV = n())
# 
# gp4_so <- get_metadata(gp4_so) %>% 
#   select(cell, UPN) %>% 
#   left_join(hmm_pred, by = "UPN") %>% 
#   column_to_rownames("cell") %>% 
#   AddMetaData(gp4_so, .)
# 
# plot_umap(gp4_so, c("methyl_nCNV", "tx_nCNV"))
```


## Other

```{r}
a <- data.table::fread("inferCNV_v2/shh/shh/infercnv.observations.txt") %>% 
  as.data.frame()
rownames(a) <- a$V1
a[, 1] <- NULL
a <- as.matrix(a)

range02 <- function(x, newMin, newMax){ (x - min(x))/(max(x)-min(x)) * (newMax - newMin) + newMin }

cin_result <- apply(a,
      2, 
      function(x){
        x <- range02(x, -1, 1)
        sum(x^2)
        #sum(abs(x - 1)) / length(x)
      })

cin_result <- data.frame(cell = names(cin_result),
                         cin = cin_result,
                         stringsAsFactors = FALSE)

mdata <- get_metadata(sos$SHH)

mdata <- inner_join(mdata, cin_result, by = "cell")

mdata %>% 
  ggplot(aes(UPN, cin)) + 
  geom_violin()
```

```{r}
shh_so <- qread(file.path("objects", "shh", "shh.qs"), nthreads = 4)
Idents(shh_so) <- "UPN"
avg_expr <- AverageExpression(shh_so)$RNA

a <- infercnv::add_to_seurat(infercnv_output_path = "inferCNV_v2/shh/shh")


hmm_pred <- read.table("inferCNV_v2/shh/shh/map_metadata_from_infercnv.txt") %>% 
  as.matrix()
hmm_pred <- t(hmm_pred[colnames(shh_so), ])

shh_so[["CNV"]] <- CreateAssayObject(data = hmm_pred)


shh_so <- get_metadata(shh_so) %>% 
  select(cell, UPN) %>% 
  left_join(methyl_cnvs, by = "UPN") %>% 
  column_to_rownames("cell") %>% 
  AddMetaData(shh_so, .)

hmm_pred <- read_tsv("inferCNV_v2/shh/shh/HMM_CNV_predictions.HMMi6.hmm_mode-samples.Pnorm_0.5.pred_cnv_regions.dat") %>% 
  separate(cell_group_name, c("id1", "id2"), sep = "\\.") %>% 
  mutate(UPN = str_remove(id1, "malignant_")) %>% 
  select(-id1, -id2) %>% 
  group_by(UPN) %>% 
  summarize(tx_nCNV = n())

shh_so <- get_metadata(shh_so) %>% 
  select(cell, UPN) %>% 
  left_join(hmm_pred, by = "UPN") %>% 
  column_to_rownames("cell") %>% 
  AddMetaData(shh_so, .)

plot_umap(shh_so, c("methyl_nCNV", "tx_nCNV"))
```

```{r}
gp4_so <- qread(file.path("objects", "gp34", "gp4_only.qs"), nthreads = 4)
Idents(gp4_so) <- "UPN"
avg_expr <- AverageExpression(gp4_so)$RNA

a <- infercnv::add_to_seurat(infercnv_output_path = "inferCNV_v2/gp4/gp4")

hmm_pred <- read.table("inferCNV_v2/gp4/gp4/map_metadata_from_infercnv.txt") %>% 
  as.matrix()
hmm_pred <- t(hmm_pred[colnames(gp4_so), ])

gp4_so[["CNV"]] <- CreateAssayObject(data = hmm_pred)


gp4_so <- get_metadata(gp4_so) %>% 
  select(cell, UPN) %>% 
  left_join(methyl_cnvs, by = "UPN") %>% 
  column_to_rownames("cell") %>% 
  AddMetaData(gp4_so, .)

hmm_pred <- read_tsv("inferCNV_v2/gp4/gp4/HMM_CNV_predictions.HMMi6.hmm_mode-samples.Pnorm_0.5.pred_cnv_regions.dat") %>% 
  separate(cell_group_name, c("id1", "id2"), sep = "\\.") %>% 
  mutate(UPN = str_remove(id1, "malignant_")) %>% 
  select(-id1, -id2) %>% 
  group_by(UPN) %>% 
  summarize(tx_nCNV = n())

gp4_so <- get_metadata(gp4_so) %>% 
  select(cell, UPN) %>% 
  left_join(hmm_pred, by = "UPN") %>% 
  column_to_rownames("cell") %>% 
  AddMetaData(gp4_so, .)

plot_umap(gp4_so, c("methyl_nCNV", "tx_nCNV"))
```


```{r}
gp3_so <- qread(file.path("objects", "gp34", "gp3_only.qs"), nthreads = 4)
Idents(gp3_so) <- "UPN"
avg_expr <- AverageExpression(gp3_so)$RNA

a <- infercnv::add_to_seurat(infercnv_output_path = "inferCNV_v2/gp3/gp3")

hmm_pred <- read.table("inferCNV_v2/gp3/gp3/map_metadata_from_infercnv.txt") %>% 
  as.matrix()
hmm_pred <- t(hmm_pred[colnames(gp3_so), ])

gp3_so[["CNV"]] <- CreateAssayObject(data = hmm_pred)


gp3_so <- get_metadata(gp3_so) %>% 
  select(cell, UPN) %>% 
  left_join(methyl_cnvs, by = "UPN") %>% 
  column_to_rownames("cell") %>% 
  AddMetaData(gp3_so, .)

hmm_pred <- read_tsv("inferCNV_v2/gp3/gp3/HMM_CNV_predictions.HMMi6.hmm_mode-samples.Pnorm_0.5.pred_cnv_regions.dat") %>% 
  separate(cell_group_name, c("id1", "id2"), sep = "\\.") %>% 
  mutate(UPN = str_remove(id1, "malignant_")) %>% 
  select(-id1, -id2) %>% 
  group_by(UPN) %>% 
  summarize(tx_nCNV = n())

gp3_so <- get_metadata(gp3_so) %>% 
  select(cell, UPN) %>% 
  left_join(hmm_pred, by = "UPN") %>% 
  column_to_rownames("cell") %>% 
  AddMetaData(gp3_so, .)

plot_umap(gp3_so, c("methyl_nCNV", "tx_nCNV"))
```

```{r}
plot_umap(sos[[2]], c("UPN", "coarse_cell_type"))

sos[[2]] <- infercnv::add_to_seurat(seurat_obj = sos[[2]], 
                               infercnv_output_path = "inferCNV_v2/gp3/gp3")

a <- read.table("inferCNV_v2/gp3/gp3/infercnv.19_HMM_predHMMi6.rand_trees.hmm_mode-subclusters.Pnorm_0.5.repr_intensities.observations.txt")

colnames(a) <- str_remove(colnames(a), "^X")

cells_to_examine <- intersect(colnames(a), colnames(sos[[2]]))

to_plot <- data.frame(x = t(a["GNG4", cells_to_examine])
                      

to_plot <- cbind(t(a["GNG4", cells_to_examine]),
                 y = sos[[2]]@assays$RNA@data["GNG4", cells_to_examine]) %>% 
  as.data.frame()

ggplot(z, aes(x = as.factor(GNG4), y = res)) + geom_boxplot()
```


```{r}
a <- read.table("inferCNV_v2/gp4/gp4/expr.infercnv.dat")

colnames(a) <- str_remove(colnames(a), "^X")

cells_to_examine <- intersect(colnames(a), colnames(sos[[3]]))

# calculate residuals from lm (expr ~ CNV)

to_test <- rownames(a)

cnv_res <- t(a[to_test, cells_to_examine])
expr_res <- Matrix::t(sos[[3]]@assays$RNA@data[to_test, cells_to_examine])

res_mat <- matrix(ncol = length(to_test), 
                  nrow = nrow(expr_res))
for(i in seq_along(to_test)){
  gene <- to_test[i]
  
  res_mat[, i] <- lm(expr_res[, gene] ~ cnv_res[, gene]) %>% 
    residuals(.) 
}

colnames(res_mat) <- to_test
rownames(res_mat) <- rownames(expr_res)

scaled_mat <- t(scale(res_mat))

so_small <- sos[[3]][rownames(scaled_mat), colnames(scaled_mat)]

so_small <- FindVariableFeatures(so_small) %>% 
  ScaleData(.) %>% 
  RunPCA() %>% 
  RunUMAP(dims = 1:30)

so_small <- SetAssayData(so_small, slot = "scale.data", new.data = scaled_mat, assay = "RNA") 
so_small <- RunPCA(so_small) %>% 
  RunUMAP(dims = 1:30, reduction.name = "res_umap")

plot_umap(so_small, "UPN")
plot_feature(so_small, "UPN", 
             embedding = "res_umap")
```