---
title: "Compare to developmental dataset"
author: "Kent Riemondy RBI"
date: "`r Sys.Date()`"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "html") })
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: cosmo
    highlight: tango
    fig_caption: true
    code_folding: hide
    df_print: paged
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libs}
source(here::here("R/utils.R"))
library(ComplexHeatmap)
library(presto)
library(openxlsx)
library(gprofiler2)
library(data.table)
library(minfi)
library(conumee)
rmd_dir <- "developmental_comparison"

fig_dir <- file.path(fig_dir, rmd_dir)

walk(fig_dir,
     dir.create, 
     showWarnings = FALSE)
```

```{r}

so_fns <- file.path(obj_dir, c(
  file.path("shh", "shh.qs"),
  file.path("gp34", "gp3_only.qs"),
  file.path("gp34", "gp4_only.qs")))
                    
clusters <- c(
  "tumor_subpopulations_shh",
  "tumor_subpopulations_gp3",
  "tumor_subpopulations_gp4"
)

sos <- map(so_fns, 
            ~qread(.x, nthreads = 4, use_alt_rep = FALSE))

names(sos) <- c("SHH", "GP3", "GP4")

```

## Compare cluster markers to cell atlas {.tabset}

The traditional way of assigning cell types to clusters is by manual inspection by an expert with familiarity with the tissue/sample/system. This is laborious, error prone, and becomingly increasingly unnecessary as the # of single cell datasets has increased.

The [primary cell atlas](http://biogps.org/dataset/BDS_00013/primary-cell-atlas/) will be used as a reference dataset to help define cell types from the cluster markers. Additionally a single cell rna-seq  [dataset](http://science.sciencemag.org/content/358/6368/1318.long) of human cortex development will be queried ("cortex-development"). As well as a dataset of [developing human cerebellum](https://www.biorxiv.org/content/10.1101/2020.06.30.174391v1) ("cerebellum-development"). 

We will use the `clustifyr` package written by RBI fellow Rui Fu, which will correlate the gene signatures in each cluster to the reference data.

### Refined clusters

First the refined clustering from the unaligned data will be compared. 

```{r, fig.width=12, fig.height=16, message=FALSE, warning=FALSE, fig.cap="Heatmap showing the correlation coefficients of each cluster against various cell types. The cell types from the human brain datasets are suffixed with cortex-development or cerebellum-development, otherwise the cells are from the primary cell atlas."}
library(clustifyr)
library(clustifyrdatahub)
ref_primary_cell_atlas <- readRDS("~/Projects/mulcahy-levy//dbases/human_primary_cell_atlas.rds")
ref_cerebellar <- readRDS("~/Projects/sc_repos/medulloblast/docs/aldinger_et_al_biorxiv_2020/avg_expr.rds")
ref_cortex_dev_local <- ref_cortex_dev() 

colnames(ref_cortex_dev_local) <- str_c(colnames(ref_cortex_dev_local),
                                        ":cortex-development")
colnames(ref_cerebellar) <- str_replace(colnames(ref_cerebellar), " ", "-") %>%
  str_remove("[0-9]{2}-") %>% 
  str_c(":cerebellar-development")

to_keep <- intersect(rownames(ref_primary_cell_atlas),
                     rownames(ref_cortex_dev_local)) %>% 
  intersect(rownames(ref_cerebellar))
          
ref <- cbind(ref_primary_cell_atlas[to_keep, ],
             ref_cortex_dev_local[to_keep, ],
             ref_cerebellar[to_keep, ])
```

```{r}
map2(sos, clusters,
    function(so, cluster){
res <- clustify(so,
                ref_mat = ref,
                cluster_col = cluster, 
                obj_out = FALSE)

hmap_mat <- t(res) %>% 
  .[rowSums(. > 0.5) > 0, ]

hmap <- Heatmap(hmap_mat,
        name = "Spearman\ncorrelation",
        row_title = "Reference cell types",
        row_names_gp = gpar(fontsize = 6),
        column_title = cluster,
        column_title_side = "bottom",
        col = viridis::magma(256))

pdf(file.path(fig_dir, str_c(cluster, "_clustifyr_cell_type.pdf")),
    width = 9,
    height = 16)
draw(hmap)
null <- dev.off()

hmap

calls <- cor_to_call(res, threshold = 0.5) %>% 
  dplyr::select(cluster, 
         called_cell_type = type, 
         correlation_value = r)

calls
})
```

### unaligned data
```{r}
so$refined_cell_type <- left_join(get_metadata(so), 
          unaligned_calls,
          by = c("refined_clusters" = "cluster")) %>% 
  select(cell, called_cell_type) %>% 
  pull(called_cell_type)

plot_umap(so, c("refined_cell_type", "UPN")) %>% 
  plot_grid(plotlist = ., nrow = 1, ncol = 2)
```

