---
title: "Marker overlap between CNVs"
author: "Kent Riemondy RBI"
date: "1/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libs, message=FALSE, warning=FALSE, echo=FALSE}
source("../R/utils.R")
library(future)

plan("multicore", workers = 7)
options(future.globals.maxSize = 2 * 1024 ^ 3)

proj_dir <- here()
data_dir <- file.path(proj_dir, "data", "cellranger", "results")
doc_dir <- file.path(proj_dir, "docs")

fig_dir <- "figs"
mkrs_dir <- "markers"
tbls_dir <- "tables"
xcel_dir <- file.path(mkrs_dir, "xlsx")
walk(c(fig_dir, mkrs_dir, tbls_dir, xcel_dir), 
     dir.create, 
     showWarnings = F)
```

## Strategy

Tumors cells tend to cluster into largely patient specific clusters. However normal cells (i.e. immune and astrocytes) tend to cluster with cells from multiple patients. This argues that the heterogeneity present in each patient is not simply an artifact from xpermental processing, but insteads reflects intrinsic heterogeneity. This hetergeneity could be driven by enviromental factors (tumor location, aggressivness, etc), or tumor genietics. Each patients CNV profile is hetergeneous, suggesting that CNV could driven expression variation, providing an explanation of the hetergeneous expression profiles.

To assess this possibility, inferCNV was run to identify CNVs. The genes in these CNV regions will be compared to markers of each patients. Need to ensure that markers that are down regulated are enriched in CNV regions that are lost and vice versa.

Here is the coding for the inferCNV hmm results:
0x: complete loss
0.5x: loss of one copy
1x: neutral
1.5x: addition of one copy
2x: addition of two copies
3x: essentially a placeholder for >2x copies but modeled as 3x.


```{r}
mkrs <- read_tsv(file.path(mkrs_dir, 
                           "upn_markers_neoplastic_pos_and_neg.tsv"),
                 col_types = c("ccnnnnnnnn")) %>% 
  filter(padj < 0.01,
         pct_in > 0.10,
         abs(logFC) > log(1.25))

subgroups <- c("shh", "gp34", "gp3", "gp4")
names(subgroups) <- c("shh", "gp34", "gp3", "gp4")
cnv_reports <- map_dfr(subgroups,
                   ~read_tsv(file.path("inferCNV",
                                       .x, 
                                      .x,
                   "HMM_CNV_predictions.HMMi6.hmm_mode-samples.Pnorm_0.5.pred_cnv_genes.dat")),
                   .id = "subgroup") 

cnv_reports <- mutate(cnv_reports, 
                      upn = str_split(cell_group_name,
                                      "\\.", 
                                      simplify = T)[, 1] %>%
                        str_remove("malignant_"),
                      change = case_when(
                        subgroup %in% c("gp34", "shh") & state < 1 ~ "loss",
                        !subgroup %in% c("gp34", "shh") & state < 3 ~ "loss",
                        subgroup %in% c("gp34", "shh") & state > 1 ~ "gain",
                        !subgroup %in% c("gp34", "shh") & state > 3 ~ "gain",
                        TRUE ~ "neutral"
                                      )) 


# get list of gnes up and down per UPN
mkr_gene_list <- split(mkrs, mkrs$group) %>% 
  map(~list(loss = filter(.x, logFC < 0) %>% pull(feature), 
            gain = filter(.x, logFC > 0) %>% pull(feature) ))


cnv_gene_list <- split(cnv_reports, cnv_reports$upn) %>%
   map(~list(loss = filter(.x, change == "loss") %>% pull(gene), 
            gain = filter(.x, change == "gain") %>% pull(gene)))

upns <- intersect(names(mkr_gene_list), names(cnv_gene_list))
```


```{r}

calc_enrichment <- function(sample_success,
                            pop_success,
                            pop_size,
                            sample_size){
  x <- sample_success - 1
  m <- pop_success 
  n <- pop_size
  k <- sample_size
  1-phyper(x, m, n, k)
}

#get some idea of a good background number

so <- readRDS("objects/all_neoplastic.rds")

n_featues_per_patient <- map_int(upns, 
                             function(upn){
                               tmp <- subset(so, cells = rownames(so@meta.data)[so$orig.ident == upn])
                               sum(Matrix::rowSums(GetAssayData(tmp, "counts")) > 0)
                             })

gene_pop_size <- median(n_featues_per_patient)

res <- map_dfr(upns, 
    function(upn){
      x <- mkr_gene_list[[upn]]$loss
      y <- cnv_gene_list[[upn]]$loss
      pop_success_loss <- length(y)
      sample_success_loss <- length(intersect(x, y))
      sample_size_loss <- length(x)
      
      x <- mkr_gene_list[[upn]]$gain
      y <- cnv_gene_list[[upn]]$gain
      pop_success_gain <- length(y)
      sample_success_gain <- length(intersect(x, y))
      sample_size_gain <- length(x)
      
      sample_success <- sample_success_gain + sample_success_loss
      pop_success <- pop_success_gain + pop_success_loss
      sample_size <- sample_size_gain + sample_size_loss
      
      pop_size <- gene_pop_size
      tibble(upn = upn,
             n_shared = sample_success,
             n_cnv_genes = pop_success,
             n_mrks = sample_size,
             pval = 1- phyper(sample_success - 1,
                             pop_success, 
                             pop_size, 
                             sample_size))
    }) 

res <- res %>% 
  mutate(prop_markers_in_cnv = n_shared / n_mrks,
         prop_genes_in_cnv_in_genome = n_cnv_genes / gene_pop_size,
         log2_effect_size = log2(prop_markers_in_cnv / prop_genes_in_cnv_in_genome))

res
```


```{r}
mdata <- so@meta.data %>% 
  as_tibble() %>% 
  select(UPN, subgroup) %>%
  unique() 

plt_dat <- res %>% 
  left_join(mdata, by = c("upn" = "UPN")) %>% 
  select(upn, 
         subgroup,
         `Marker genes\nin CNVs` = prop_markers_in_cnv, 
         `CNV genes\nin transcriptome` = prop_genes_in_cnv_in_genome) %>% 
  pivot_longer(cols = -c(upn, subgroup)) 

p <- ggplot(plt_dat, aes(upn, value)) +
  geom_col(aes(fill = name), position = position_dodge2()) +
  labs(x = "",
       y = "Proportion") +
  facet_grid(subgroup ~ ., space = "free", scales = "free") + 
  scale_fill_manual(name = "",
                    values = palette_OkabeIto) + 
  coord_flip() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        legend.position = "top",
        strip.text = element_text(size = 6))

save_plot("figs/cnv_marker_enrichment.pdf", p, base_asp = 0.75, base_height = 6.5)
```
For most tumors there is a signficant overlap between marker genes and CNV regions. 1202 and 1224 have non-significant overlap. 


# Examine subclones

For some tumors it appears that there are some subclones based on CNV. For this anslysis I will split each tumor into 2 subclones based on HClustering on the CNV results. Then genexpression between theses splits will be evaluated to see if there are DE genes. 

```{r}
library(ape)
library(dendextend)
library(scbp)
so <- readRDS("objects/gp34.rds")

subgroups <- c(
  "shh",
  "gp3",
  "gp34",
  "gp4"
)

trees <- map(subgroups,
             ~read.tree(
               file.path("inferCNV",
             .x, .x,  "infercnv.observations_dendrogram.txt"))) 

# merge dendograms, try to figure out reasonable split
ds <- map(trees, 
    ~map(.x, as.dendrogram)) %>% 
  map(., 
      ~Reduce(function(x, y){merge(x,y)}, .x))

z <- labels(ds[[1]]) %>% str_split("_", simplify = TRUE) %>% .[, 1] %>% as.numeric(as.factor(.))

a <- color_branches(ds[[1]], clusters = z) 
plot(a)

map(trees, 
    ~map(.x, ~dendextend::cutree(.x, k = 2)) %>% unlist()) %>%
  unlist() -> a

so$cnv <- a[colnames(so)]
so$cnv <- ifelse(is.na(so$cnv), -1,so$cnv)
```

```{r}

to_plot <- c(
  "cnv",
  "UPN",
  "nCount_RNA",
  "subgroup"
)

map2(to_plot, 
    c("by_feature",
      "random",
      "random",
      "random"),
    ~plot_feature(so, 
               .x, 
               pt_alpha = 0.7, 
               sorted = .y,
               embedding = "umap")) %>% 
  plot_grid(plotlist = .)
```

```{r}
plot_violin(get_metadata(so), "cnv", "nCount_RNA")

```