---
title: "Immune populations"
author: "Kent Riemondy RBI"
date: "`r Sys.Date()`"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "html") })
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: cosmo
    highlight: tango
    fig_caption: true
    code_folding: hide
    df_print: paged
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libs, message=FALSE, warning=FALSE, echo=FALSE}
source(here::here("/R/utils.R"))

rmd_dir <- "immune"

fig_dir <- file.path(fig_dir, rmd_dir)
mkrs_dir <- file.path(mkrs_dir, rmd_dir)
tbls_dir <- file.path(tbls_dir, rmd_dir)
obj_dir <- file.path(obj_dir, rmd_dir)

walk(c(fig_dir, mkrs_dir, tbls_dir, obj_dir),
     dir.create, showWarnings = F)


seed_value <- 20200204
```

```{r}
so <- qread(file.path("objects", "preprocess", "so.qs"))
```


```{r}
immune_cells <- colnames(so)[so$coarse_cell_type %in% c(
                               "lymphocytes",
                               "macrophage_monocytes")]

so <- so[, immune_cells]
so <- subset(so, subset = subgroup != "WNT")

so <- FindVariableFeatures(so, 
                           selection.method = "vst", 
                           nfeatures = 3000, 
                           verbose = FALSE)

so <- ScaleData(so, 
                vars.to.regress = c("percent.mt", "nCount_RNA"), 
                verbose = TRUE)

so <- RunPCA(so, npcs = 40, verbose = FALSE, seed.use = seed_value)

so <- FindNeighbors(so,
                    reduction = "pca",
                    dims = 1:30,
                    k.param = 20)

so <- RunUMAP(so, graph = "RNA_snn", seed.use = seed_value)

so <- FindClusters(so, 
                   random.seed = seed_value,
                   resolution = c(0.4, 0.3, 0.25, 0.2, 0.1, 0.05))
```


```{r}
to_plot <- c(
  "CD79A",
  "CD3G",
  "CCL3",
  "S100A9",
  "FCER1A",
  "NAV3",
  "FOXP3",
  "MKI67",
  "SEZ6L2"
)

plot_umap(so, to_plot) %>% 
  plot_grid(plotlist = ., 
            nrow = 3,
            ncol = 3)
```

```{r}
plot_umap(so, 
          str_subset(colnames(so@meta.data), "res")) %>% 
  plot_grid(plotlist = ., 
            nrow = 2,
            ncol = 3)
```


```{r}
so$fine_immune_clusters <- so$RNA_snn_res.0.3

clusters <- so$fine_immune_clusters

p <- plot_umap(so, "fine_immune_clusters", label_text = TRUE)
p
save_plot(file.path(fig_dir, "umap_by_clusters.pdf"),
                     p, base_asp = 1.4)

p <- plot_umap(so, "subgroup")
p
save_plot(file.path(fig_dir, "umap_by_subgroup.pdf"), p, base_asp = 1.2)

p <- plot_umap(so, "UPN")
p
save_plot(file.path(fig_dir, "umap_by_upn.pdf"), p, base_asp = 1.2)
```

## Remove patient specific cluster


````{r}
so <- calc_diversity(so, sample_id = "orig.ident", group_id = "fine_immune_clusters")

plot_umap(so, "entropy")
```

Remove cluster 3 and reembed

```{r}
so <- subset(so, subset = fine_immune_clusters != "3")

so <- FindVariableFeatures(so, 
                           selection.method = "vst", 
                           nfeatures = 3000, 
                           verbose = FALSE)

so <- ScaleData(so, 
                vars.to.regress = c("percent.mt", "nCount_RNA"), 
                verbose = TRUE)

so <- RunPCA(so, npcs = 40, verbose = FALSE, seed.use = seed_value)

so <- FindNeighbors(so,
                    reduction = "pca",
                    dims = 1:30,
                    k.param = 20)

#so <- RunUMAP(so, dims = 1:30, graph = "RNA_snn", seed.use = seed_value)
so <- RunUMAP(so, dims = 1:30, seed.use = seed_value)

so <- FindClusters(so, 
                   random.seed = seed_value,
                   resolution = c(0.4, 0.3, 0.25, 0.2, 0.1, 0.05))
```


```{r}
to_plot <- c(
  "CD79A",
  "CD3G",
  "CCL3",
  "S100A9",
  "FCER1A",
  "NAV3",
  "FOXP3",
  "MKI67",
  "SEZ6L2"
)

p <- plot_umap(so, to_plot) %>% 
  plot_grid(plotlist = ., 
            nrow = 3,
            ncol = 3)

p
save_plot(file.path(fig_dir, "umap_marker_summary.pdf"), 
          p,
          nrow = 3, 
          ncol = 3,
          base_asp = 1.2)

```

```{r}
plot_umap(so, 
          str_subset(colnames(so@meta.data), "res")) %>% 
  plot_grid(plotlist = ., 
            nrow = 2,
            ncol = 3)
```


```{r}
so$fine_immune_clusters <- so$RNA_snn_res.0.3

clusters <- so$fine_immune_clusters

p <- plot_umap(so, "fine_immune_clusters", label_text = TRUE)
p
save_plot(file.path(fig_dir, "umap_by_clusters.pdf"),
                     p, base_asp = 1.4)

p <- plot_umap(so, "subgroup")
p
save_plot(file.path(fig_dir, "umap_by_subgroup.pdf"), p, base_asp = 1.2)

p <- plot_umap(so, "UPN")
p
save_plot(file.path(fig_dir, "umap_by_upn.pdf"), p, base_asp = 1.2)
```

## Annotate cell types

Using Human primary cell atlas https://www.ncbi.nlm.nih.gov/pubmed/24053356


```{r, fig.height=20, fig.width=8}
library(clustifyr)
hpca_ref <- readRDS("../dbases/human_primary_cell_atlas.rds")

res <- clustify(
  so,
  hpca_ref, 
  cluster_col = "fine_immune_clusters",
  query_genes = so@assays$RNA@var.features,
  seurat_out = FALSE
)

library(ComplexHeatmap)

hmap <- Heatmap(t(res), 
                viridis::viridis(256), 
                "Spearman\ncorrelation",
                row_title = "Cell types from HCPA",
                column_title = "Clusters")
hmap

pdf(file.path(fig_dir, "hmap_hpca.pdf"),
    width = 8, height = 22)
  print(hmap)
dev.off()
```
## add in cluster diversity


```{r}
so <- calc_diversity(so, sample_id = "UPN", group_id = "fine_immune_clusters")
so$cdiversity <- so$entropy
so$entropy <- NULL
p <- plot_umap(so, "cdiversity")
p
save_plot(file.path(fig_dir, "umap_by_cluster_diversity.pdf"), p, base_asp = 1.2)
```

Cluster 
```{r}
mdata <- so@meta.data %>% 
  rownames_to_column("cell")
ctypes  <- call_to_metadata(cor_to_call(res, 
                             metadata = mdata, 
                             cluster_col = "fine_immune_clusters", threshold = 0.5),
                 metadata = mdata,
                 cluster_col = "fine_immune_clusters") %>% 
  mutate(type = ifelse(type == "r<0.5, unassigned",
                       "unknown",
                       type)) %>% 
  pull(type)

# keep broad cell type definition
so$hpca_cell_types <- ctypes %>% 
  str_split(., ":") %>% 
  map_chr(~str_c(.x[1:(length(.x) - 1)], collapse = ":"))

so$hpca_fine_cell_types <- ctypes 

p <- plot_umap(so, "hpca_fine_cell_types")
p
save_plot(file.path(fig_dir, "umap_by_coarse_cell_type.pdf"), p, base_asp = 1.2)

p <- plot_umap(so, "hpca_cell_types")
p
save_plot(file.path(fig_dir, "umap_by_fine_cell_type.pdf"), p, base_asp = 1.2)
```




## Immune cluster markers

```{r}
library(presto)

Idents(so) <- "fine_immune_clusters"

mkrs <- wilcoxauc(so, "fine_immune_clusters")

mkrs <- filter(mkrs, logFC > 0, padj < 0.05, pct_in > 0.10) %>% 
  group_by(group) %>% 
  arrange(padj, desc(logFC), .by_group = TRUE)

mkrs %>% 
  write_tsv(file.path(mkrs_dir, "immune_cluster_markers.tsv"))

mkrs %>% 
  ungroup() %>% 
  split(., .$group) %>% 
  write_markers_xlsx(.,
                     file.path(tbls_dir, "immune_cluster_markers.xlsx"))

```

## Cell type composition by subgroup


```{r}
p <- plot_features_split(so, "hpca_cell_types", "subgroup")

p
save_plot(file.path(fig_dir, 
                    "umap_by_fine_cell_types_split_by_subgroup.pdf"), p,
          nrow = 2,
          ncol = 2,
          base_asp = 1.2)
```

```{r}
library(MASC)
library(lme4)

mdata <- so@meta.data

#compare shh to grp3/4
mdata$subgroup_stat_id <- ifelse(mdata$subgroup == "SHH",
                                 "SHH",
                                 "Gp3and4") %>% 
  as.factor()

composition_stats <- MASC:::MASC(data =  mdata, 
            cluster = mdata$fine_immune_clusters, 
            contrast = "subgroup_stat_id",
            random_effects = "UPN") 


composition_stats <- mutate(composition_stats, 
                            log2_OR = log2(subgroup_stat_idSHH.OR),
                            log2_OR_lower = log2(subgroup_stat_idSHH.OR.95pct.ci.lower),
                            log2_OR_upper = log2(subgroup_stat_idSHH.OR.95pct.ci.upper),
                            sig = case_when(
                              model.pvalue < 0.0001 ~ "***",
                              model.pvalue < 0.001 ~ "**",
                              model.pvalue < 0.05 ~ "*",
                              TRUE ~ ""))
                            
composition_stats <- filter(composition_stats, is.finite(subgroup_stat_idSHH.OR.95pct.ci.upper))


p <- ggplot(composition_stats, aes(cluster, log2_OR)) + 
  geom_pointrange(aes(color = -log10(model.pvalue),
                      ymin = log2_OR_lower,
                      ymax = log2_OR_upper)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_text(aes(label = sig), nudge_y = 1) + 
  scale_color_gradientn(colours = rev(viridis::viridis(256))) +
  coord_flip() +
  labs(y = "Odds ratio (log2)",
       color = "-log10(p)",
       subtitle = "SHH vs. Group3+4")

p

save_plot(file.path(fig_dir, "composition_shift_odds.pdf"), 
          p,
          base_asp = 1.2)
```

```{r}
p <- plot_umap(so, "fine_immune_clusters", group = "subgroup")

p

save_plot(file.path(fig_dir, "umap_split_by_cluster.pdf"), 
          p,
          nrow = 2,
          ncol = 2,
          base_asp = 1)

```

## Additional annotation


```{r}

immune_new_long_ids <- c(
  "8" = "B-cells (CD79A+, IGHM+)",
  "10" = "Treg (FOXP3+, CTLA4+)",
  "9" = "Proliferative (MKI67+, TOP2A+)",
  "6" = "Microglia (P2RY12+ TMEM119+)",
  "0" = "T-cells (CD3G+, IL32+)",
  "7" = "NK-cells (GNLY+, NKG7+)",
  "5" = "Neutrophil/Monocyte (S100A9+, THBS1+)",
  "4" = "Dendritic-cell-like (CD1C+,FCER1A+)",
  "2" = "TAM CytoHi+ (CCL3+, SPP1+)",
  "1" = "TAM M1/M2 (APOE+, TYROBP+)",
  "3" = "TAM M1/M2 (MRC1+ CD163+)"
)
so@meta.data$new_cell_type_ids <- immune_new_long_ids[as.character(so$fine_immune_clusters)]
   

so@meta.data$new_cell_type_ids_short <- str_split(so$new_cell_type_ids, 
                                                  " ",
                                                  simplify = TRUE) %>% 
  .[, 1]
   
plot_umap(so, "new_cell_type_ids")
plot_umap(so, "new_cell_type_ids_short")
```

## Immune cluster markers

```{r}
library(presto)

Idents(so) <- "new_cell_type_ids"

full_mkrs <- wilcoxauc(so, "new_cell_type_ids")

mkrs <- filter(full_mkrs, logFC > 0, padj < 0.05, pct_in > 0.10) %>% 
  group_by(group) %>% 
  arrange(padj, desc(logFC), .by_group = TRUE)

mkrs %>% 
  write_tsv(file.path(mkrs_dir, "immune_cell_type_markers.tsv"))

mkrs %>% 
  ungroup() %>% 
  split(., .$group) %>% 
  write_markers_xlsx(.,
                     file.path(tbls_dir, "immune_cell_type_markers.xlsx"))

full_mkrs %>% 
  write_tsv(file.path(mkrs_dir, "immune_cell_type_full_results.tsv"))


```

### Each cell type

```{r}
library(gprofiler2)
library(ggrepel)
mkrs <- read_tsv(file.path(mkrs_dir, 
                    "immune_cell_type_markers.tsv")) %>% 
  filter(logFC > 0, padj < 0.05, pct_in > 0.10) %>% 
  group_by(group) %>% 
  arrange(padj, desc(logFC), .by_group = TRUE) %>% 
  slice(1:500)

res  <- map(unique(mkrs$group), function(x){
  clust_mkrs <- filter(mkrs, group == x) %>% 
    pull(feature) %>% 
    .[1:500]
  go_res <- gost(clust_mkrs, 
                 ordered_query = TRUE,
                 sources = c("GO", 
                             "KEGG", 
                             "REAC", 
                             "TF"))
  p <- gostplot(go_res, 
                interactive = FALSE)
  
  
  text_data <- group_by(p$data, 
                        source) %>% 
    slice(1:2)
  
  p + 
    geom_text_repel(data = text_data, aes(label = term_name))

    go_res[["result"]] %>% 
    mutate(parents = unlist(map(parents, ~str_c(.x, collapse = ","))),
           source = str_replace_all(source, "[[:punct:]]", " ")) %>% 
    select(-significant, -query)
})

names(res) <- str_c("cluster_", unique(mkrs$group))

openxlsx::write.xlsx(res,
                     file.path(tbls_dir,"immune_cell_type_goterms.xlsx"))
  
```



## Cell proportions

```{r}
p <- plot_cell_proportions(so, 
                      sample_id = "UPN",
                      group_id = "new_cell_type_ids",
                      facet_by = "subgroup") +
  theme(strip.text = element_text(angle = 45))
p
```


## Average Expression per cluster

```{r}
write_avg_expr(so,  
               "new_cell_type_ids",
               file.path(tbls_dir, 
                         "new_cell_type_id_avgexpr.tsv.gz"))

cell_count_mat <- get_cell_count_matrix(so, "UPN", "new_cell_type_ids")

cell_count_mat[is.na(cell_count_mat)] <- 0L

cell_count_mat %>% 
  tibble::rownames_to_column("UPN") %>% 
write_csv(file.path(tbls_dir, 
                         "immune_cell_type_cell_counts.csv"))


full_marker_matrix <- full_mkrs %>% 
  top_marker_matrix(.) 
  
topn_marker_matrix <- full_mkrs %>% 
  top_marker_matrix(n = 200) 

out <- list(avg_expression = read_tsv(file.path(tbls_dir, "new_cell_type_id_avgexpr.tsv.gz")),
     logFC_all = full_marker_matrix,
     top_200_logFC = topn_marker_matrix)

openxlsx::write.xlsx(out, 
                     file.path(tbls_dir, "immune_cell_type_expression_logfc_summaries.xlsx"))
```


```{r}
GetAssayData(so, "counts") %>% 
  as.matrix(.) %>% 
  as.data.frame() %>% 
  rownames_to_column("gene") %>% 
  data.table::fwrite(., file.path(tbls_dir, "immune_count_matrix.tsv.gz"),
       sep = "\t",
       compress = "gzip")
  
# expression matrix filtered at least 25 cells express the gene
expr_mat <- GetAssayData(so, "data")
expr_mat <- expr_mat[Matrix::rowSums(expr_mat > 0) >= 25, ]

expr_mat %>% 
  as.matrix(.) %>% 
  as.data.frame() %>% 
  rownames_to_column("gene") %>% 
  data.table::fwrite(., file.path(tbls_dir, "immune_expr_matrix.tsv.gz"),
       sep = "\t",
       compress = "gzip")
```

## save object


```{r eval = FALSE}
qsave(so, file.path(obj_dir, "immune_so.qs"))
```

```{r, eval = FALSE}
so <- qread(file.path(obj_dir, "immune_so.qs"), use_alt_rep = FALSE )
```
## Markers within cluster

```{r}
Idents(so) <- "hpca_cell_types"

so_sub <- subset(so, 
                 subset = hpca_cell_types == "Macrophage:monocyte-derived")

mkrs <- FindMarkers(so_sub, 
                    ident.1 = "SHH", 
                    group.by = "subgroup")

mkrs <- tibble::rownames_to_column(mkrs, "gene")

write_tsv(mkrs, 
          file.path(mkrs_dir, 
                    "macrophage_markers_shh_to_others.tsv"))

mkrs <- set_xlsx_class(mkrs, "gene", "Text")
openxlsx::write.xlsx(mkrs, 
          file.path(tbls_dir, 
                    "macrophage_markers_shh_to_others.xlsx"))

```

```{r}
mkrs <- arrange(mkrs, 
                p_val_adj)
up_mkrs <- filter(mkrs, avg_logFC > 0) %>% slice(1:30)
down_mkrs <- filter(mkrs, avg_logFC < 0) %>% slice(1:30)

top_mkrs <- bind_rows(up_mkrs, down_mkrs)

library(ComplexHeatmap)

Idents(so_sub) <- "subgroup"

to_plot <- log1p(AverageExpression(so_sub)$RNA)
to_plot <- to_plot[top_mkrs$gene, ] 

to_plot <- t(scale(t(as.matrix(to_plot))))

hm <- Heatmap(to_plot, col = viridis::viridis(256),
        name = "Z-score")

pdf(file.path(fig_dir, "hmap_macrophages_differences_between_subgroups.pdf"), height = 12)
print(hm)
dev.off()


```


```{r, eval = FALSE}
so_sub <- subset(so, 
                 subset = refined_immune_clusters == "t-cells")

mkrs <- FindMarkers(so_sub, 
                    ident.1 = "GP4", 
                    ident.2 = "SHH", 
                    group.by = "subgroup")

mkrs <- tibble::rownames_to_column(mkrs, "gene")

write_tsv(mkrs, 
          file.path(mkrs_dir, 
                    "tcells_markers_between_gp4_shh.tsv"))

```

```{r, eval = FALSE}

mkrs <- arrange(mkrs, 
                p_val_adj)
up_mkrs <- filter(mkrs, avg_logFC > 0) %>% slice(1:50)
down_mkrs <- filter(mkrs, avg_logFC < 0) %>% slice(1:50)

top_mkrs <- bind_rows(up_mkrs, down_mkrs)

library(ComplexHeatmap)

Idents(so_sub) <- "subgroup"

to_plot <- log1p(AverageExpression(so_sub)$RNA)
to_plot <- to_plot[top_mkrs$gene, ] 

to_plot <- t(scale(t(as.matrix(to_plot))))

hm <- Heatmap(to_plot, col = viridis::viridis(256),
        name = "Z-score")

pdf("t-cell_differences_between_subgroups.pdf", height = 16)
print(hm)
dev.off()


xcel_out$t_cells <- mkrs

xcel_out <- map(xcel_out, 
                       ~set_xlsx_class(.x, "gene", "Text"))

openxlsx::write.xlsx(xcel_out, 
                     "markers/shh_gp4_immune_cell_differences.xlsx")
```


```{r}
to_plot <- c("refined_immune_clusters", 
"subgroup",
  "UPN")

plts <- map(to_plot, ~plot_umap(so, .x))



plt <- plot_grid(plotlist = plts, nrow = 2, ncol = 2,
                 rel_widths = c(1.75, 1.25))
save_plot("immune_population_umaps.pdf", plt, 
          nrow = 2, ncol = 2, base_asp = 1.2)


plts <- map(c("MRC1", "CCL3"), ~plot_umap(so, .x))


plt <- plot_grid(plotlist = plts, nrow = 1)
save_plot("immune_population_umaps_expr.pdf", plt, 
          nrow = 1, ncol = 2, base_asp = 1.2)

```

```{r}
sc_mdata <- so@meta.data %>% 
  rownames_to_column("cell")

new_annotation <- so@meta.data %>% 
    rownames_to_column("cell") %>% 
    group_by(hpca_cell_types, fine_immune_clusters) %>%
    summarize() %>%
    mutate(fine_cell_type_ids = paste0(hpca_cell_types, "_", "#", row_number())) %>% 
    ungroup() %>% 
  select(fine_immune_clusters, fine_cell_type_ids)

so$fine_cell_type_ids <- left_join(sc_mdata, new_annotation, by = c("fine_immune_clusters")) %>% 
  pull(fine_cell_type_ids)  

sc_mdata <- so@meta.data %>% 
  rownames_to_column("cell")

per_patient <- sc_mdata %>% 
  filter(hpca_cell_types == "Macrophage:monocyte-derived") %>% 
  group_by(UPN) %>%
  mutate(n_cells = n()) %>% 
  group_by(UPN, fine_cell_type_ids) %>% 
  summarize(n = n(),
            prop_cell_type = n / unique(n_cells),
            subgroup = unique(subgroup))

p <- ggplot(per_patient, 
       aes(UPN, prop_cell_type)) +
  geom_col(aes(fill = fine_cell_type_ids)) +
  labs(x = "Tumor ID",
       y = "Proportion of each cell type") +
  scale_fill_manual(values = discrete_palette_default, 
                    name = "") + 
  facet_grid(~subgroup, scales = "free_x", space = "free_x") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position = "top",
        strip.background = element_rect(fill = "white"))
p
```


## Gene set enrichment

### Macrophages
```{r}
library(gprofiler2)
library(ggrepel)
mkrs <- read_tsv(file.path(mkrs_dir, 
                    "macrophage_markers_shh_to_others.tsv")) %>% 
  filter( p_val_adj < 0.05)

go_res <- gost(mkrs$gene, 
               ordered_query = TRUE,
               sources = c("GO", 
                          "KEGG", 
                          "REAC", 
                          "TF"))
p <- gostplot(go_res, 
              interactive = FALSE)
              

text_data <- group_by(p$data, 
                      source) %>% 
  slice(1:2)

p + 
  geom_text_repel(data = text_data, aes(label = term_name))

go_res[["result"]] %>% 
  mutate(parents = unlist(map(parents, ~str_c(.x, collapse = ",")))) %>% 
  write_tsv(., file.path(tbls_dir, "monocyte_shh_gp4_goterms.tsv"))

go_res[["result"]] %>% 
  mutate(parents = unlist(map(parents, ~str_c(.x, collapse = ","))),
         source = str_replace_all(source, "[[:punct:]]", " ")) %>% 
  select(-significant, -query) %>% 
  split(., .$source) %>% 
  openxlsx::write.xlsx(., file.path(tbls_dir, "monocyte_shh_gp4_goterms.xlsx"))
```

## Add better cell type labels

```{r}

new_ids <- c(
  "6" = "Non-activated microglia (NA-M)",
  "1" = "Complement myeloid (Complement-M)",
  "3" = "M2-activated myeloid  (M2-M)",
  "2" = "Chemokine myeloid (Chemokine-M)",
  "4" = "Dendritic cell-like myeloid (DC-M)",
  "5" = "Neutrophil (Nt)",
  "7" = "NK cell (NK)",
  "0" = "T cell (T)",
  "8" = "B cell (B)",
  "10" = "Treg cell (Treg)",
  "9" = "Proliferative (Prolif.)"
)

new_ids <- factor(new_ids, levels = new_ids)

so$publication_cell_types <- new_ids[as.character(so$fine_immune_clusters)]

p <- plot_umap(so, "publication_cell_types", legend_title = "", sorted = "random")


embed_med_dat <- p$data %>%
  group_by(publication_cell_types) %>%
  mutate(median_x = median(`UMAP_1`),
         median_y = median(`UMAP_2`)) %>%
  mutate(new_id = ifelse(scbp:::closest_to_point(data.frame(.data[["UMAP_1"]],
                                                     .data[["UMAP_2"]]),
                                          c(unique(median_x),
                                            unique(median_y))),
                         as.character(str_match(publication_cell_types,
                                                "\\((.+)\\)") %>% .[, 2]),
                         "")
  )

# use same colors as each feature
p <- p + ggrepel::geom_text_repel(data = embed_med_dat,
                           aes_string(x = "median_x",
                                      y = "median_y",
                                      label = "new_id"),
                           size = 3,
                           segment.colour = "black",
                           force = 0.4,
                           nudge_y = 1,
                           seed = 2)
p

save_plot(file.path(fig_dir, "publication_cell_type_labels.pdf"),
          p,
          base_asp = 2.0)
```



